---
title: "Parameterized Report"
output: html_document
params:
  year:
    label: "Year(s):"
    value: 2015
    input: select
    choices: [2016, 2015, 2014]
  university:
    label: "University ID:"
    value: 222178
    input: numeric
    min: 100654
    max: 489937
---

```{r setup, include=FALSE}
library(plotly)
library(tidyverse)
library(scales)
library(treemapify)
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)

tuition <- sprintf("https://api.datausa.io/api/csv/?sort=desc&show=university&required=oos_tuition,state_tuition&sumlevel=2&year=%s&university=%d", params$year, params$university) %>% read.csv(na.strings = "None") 

admissions <- sprintf("https://api.datausa.io/api/csv/?sort=desc&show=university&required=applicants_total,admissions_total&sumlevel=2&year=%s&university=%d", params$year, params$university) %>% read.csv(na.strings = "None")

sat <- sprintf("https://api.datausa.io/api/csv/?sort=desc&display_names=1&show=university&required=sat_math_25,sat_math_75,sat_cr_25,sat_cr_75,sat_writing_25,sat_writing_75&sumlevel=2&year=%s&university=%d", params$year, params$university) %>% read.csv(na.strings = "None")

enrollment <- sprintf("https://api.datausa.io/api/csv/?sort=desc&display_names=1&show=enrollment_status&required=enrollment_status,num_enrolled&sumlevel=all&year=%s&university=%d", params$year, params$university) %>% read.csv(na.strings = "None") 

admission_gender <- sprintf("https://api.datausa.io/api/csv/?sort=desc&show=university&required=applicants_men,applicants_women,admissions_men,admissions_women&sumlevel=2&year=%s&university=%d", params$year, params$university) %>% read.csv(na.strings = "None")

enrollment_race <- sprintf("https://api.datausa.io/api/csv/?sort=desc&show=ipeds_race&required=num_enrolled&sumlevel=all&year=%s&university=%d,MAS", params$year, params$university) %>% read.csv(na.strings = "None") %>% mutate(ipeds_race_name = as.character(ipeds_race_name)) %>% arrange(num_enrolled) %>% group_by(university_name) %>% mutate(perc=num_enrolled/sum(num_enrolled))

grad_race <- sprintf("https://api.datausa.io/api/csv/?sort=desc&sumlevel=all,all&show=ipeds_race,sex&year=%s&university=%d&required=grad_rate,num_finishers", params$year, params$university) %>% read.csv(na.strings = "None") 

expenses <- sprintf("https://api.datausa.io/api/csv/?sort=desc&show=ipeds_expense&required=benefits_expense,dep_expense,interest_expense,ops_expense,other_expense,salaries_expense&sumlevel=all&year=%s&university=%d", params$year, params$university) %>% read.csv(na.strings = "None") 

```

## Report for `r unique(admissions$university_name)` for the `r params$year` year

`r unique(admissions$university_name)` has `r prettyNum(sum(enrollment$num_enrolled), big.mark = ",")` students enrolled. In `r params$year`, `r prettyNum(admissions$applicants_total, big.mark = ",")` prospective undergraduate students applied to the university. Of those students, `r prettyNum(admissions$admissions_total, big.mark = ",")` (`r round(admissions$admissions_total/admissions$applicants_total*100, digits=0)` percent) were admitted. It cost \$`r prettyNum(tuition$state_tuition, big.mark = ",")` in `r params$year` for in-state and \$`r prettyNum(tuition$oos_tuition, big.mark = ",")` for out-of-state students.

For students admitted in `r params$year`, SAT scores for the 25th and 75th percentile were `r sat$sat_cr_25` and `r sat$sat_cr_75` on critical reading and `r sat$sat_math_25` and `r sat$sat_math_75` on math.

### Undergraduate Admissions

This chart compares the `r params$year` acceptance rate of `r unique(admissions$university_name)` by gender.

```{r}
admission_gender %>% 
  group_by(year, university_name) %>% 
  summarise(accept_men=admissions_men/applicants_men,
         accept_women=admissions_women/applicants_women) %>% 
  gather(starts_with("accept"), key = "gender", value = "rate") %>% 
  separate(gender, into = c("type", "gender")) %>% 
  mutate(gender=factor(gender, levels = c("women", "men"), labels =  c("Female", "Male"))) %>% 
  ggplot() +
  geom_col(aes(x=factor(year), y=rate, fill=gender), position = position_dodge()) +
  theme(legend.position = "bottom",
        legend.title=element_blank()) +
  scale_y_continuous(name="Acceptance Rate", labels=percent) +
  scale_x_discrete(name="", expand = c(0.01,0.01)) +
  coord_flip() 
```

### Enrollment

The chart below shows the enrolled student population at `r unique(admissions$university_name)` by ethnicity in `r params$year`. This includes both full-time and part-time students as well as graduate and undergraduates.

```{r enrollment, eval=TRUE}
ggplot() +
  geom_col(data = enrollment_race, aes(x=factor(ipeds_race_name), y=perc, fill=university_name), position = position_dodge()) +
  scale_y_continuous(name="Share", labels=percent) +
  scale_x_discrete(name="") +
  coord_flip() +
  theme(legend.position = "bottom",
        legend.title=element_blank())
```

### Completions

```{r}
grad_race %>% 
  ggplot() + 
  geom_col(mapping = aes(x=ipeds_race_name, y=grad_rate, fill=sex_name), 
           position = position_dodge()) +
  theme(legend.position = "bottom",
        legend.title=element_blank()) +
  scale_y_continuous(name="Graduation Rate", labels=percent) +
  scale_x_discrete(name="") +
  coord_flip()
```


### Expenses

This tree map shows all of the primary expenses by category at `r unique(admissions$university_name)` as a share of total expenditure.

```{r, eval=(nrow(expenses)<1)} 
paste0("No data currently available for the ", params$year, " year.") 
```

```{r, eval=nrow(expenses)>0, fig.height=10, fig.width=10}
expenses %>% 
  mutate_at(.vars = vars(university_name,
                         ipeds_expense_name,
                         ipeds_expense),
            as.character) %>% 
  mutate_if(is.factor, ~as.numeric(levels(.))[.]) %>% 
  gather(contains("expense"), -ipeds_expense_name, -ipeds_expense, key = "expense", value = "value") %>% 
  separate(expense, into = c("expense", "type")) %>% 
  select(-type) %>% 
  group_by(expense, ipeds_expense_name) %>% 
  summarise(rate=sum(value, na.rm = TRUE)) %>% 
  mutate(perc=rate/sum(abs(rate)),
         expense_name = case_when(
           expense == "salaries" ~ "Salaries",
           expense == "benefits" ~ "Benefits",
           expense == "ops" ~ "Operations",
           expense == "dep" ~ "Departmental",
           expense == "interest" ~ "Interest",
           TRUE ~ "Other"
         )) %>% 
  ggplot(mapping = aes(area=rate, fill=perc, label=paste0(ipeds_expense_name, "\n", round(perc*100, 1),"%"),
                       subgroup=expense_name)) +
  geom_treemap() +
  geom_treemap_subgroup_border() +
  geom_treemap_subgroup_text(place = "bottomright", grow = TRUE, colour =
                             "white", fontface = "italic", min.size = 0) +
  geom_treemap_text(fontface = "bold", color = "white", place = "topleft", reflow = TRUE, min.size = 0, alpha = 0.9) +
  scale_fill_gradient(low="#00BFC4", high="#F8766D", labels=percent) +
  theme(legend.position = "bottom") +
  labs(fill="Share")
```


### Knitr and Rmarkdown Engines

#### Python

```{python, echo=TRUE, eval=FALSE}
x = 'hello, python world!'
print(x.split(' '))
```

#### SQL

```{r, echo=TRUE}
library(DBI)
db <- dbConnect(RSQLite::SQLite(), dbname = "sql.sqlite")
```

```{sql, connection=db, max.print = 20, eval=FALSE, echo=TRUE}
SELECT * FROM trials
```

#### CSS

```{css, echo=TRUE}
body {
  font-family: arial;
  max-width: 1200px;
  margin-left: auto;
  margin-right: auto;
  background-color: #582C83;
}

.main-container {
  max-width: 1000px;
  margin-left: auto;
  margin-right: auto;
  background-color: #fff;
}
```

